---
title: "Forensic Glass:  Recursive Feature Elimination"
output: 
  html_document:
    theme: united
    toc: yes
  html_notebook:
    theme: united
    toc: yes
---

efg | 2018-09-06

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
time.1 <- Sys.time()
```

The `caret` package has a **recursive feature elimination** function, `rle`, which is a backward selection approach.  We start with all features, remove the least important feature, and then update the model.

## Setup

```{r, comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(MASS)          # fgl data
library(caret)         # createDataParition
library(doParallel)    # registerDoParallel  
library(kableExtra)    
```

## Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
dim(rawData)
```

```{r, comment=NA}
rawData %>% head(5)
```


Relatively small "n" and class imbalance will make machine learning difficult.

## Define train and test datasets

```{r}
set.seed(71)

trainSetIndices <- createDataPartition(rawData$type, p=0.70, list=FALSE)

trainSet <- rawData[ trainSetIndices, ]
testSet  <- rawData[-trainSetIndices, ]   
```

```{r}
dim(trainSet)
```

```{r}
x <- trainSet[,-ncol(trainSet)]  # matrix for model training
y <- trainSet[, ncol(trainSet)]  # vector of training outcomes
```

## Recursive Feature Elimination (RFE)

From the [online help](http://topepo.github.io/caret/recursive-feature-elimination.html#rfe):  *There are a number of pre-defined sets of functions for several models, including: linear regression (in the object`lmFuncs`), random forests (`rfFuncs`), naive Bayes (`nbFuncs`), bagged trees (`treebagFuncs`) and functions that can be used with  caret ’s  train  function (`caretFuncs`).*

```{r}
subsets <- 1:ncol(x)

nCV      <-  3 # 3-fold cv (since N is fairly small)
nRepeats <- 50 # 50 cv repeats
```

Repeatable random numbers.  See ?refControl.

```{r}
set.seed(73)
seeds <- vector(mode = "list", length = nCV*nRepeats+1)
for (i in 1:(nCV*nRepeats)) seeds[[i]] <- sample.int(nCV*nRepeats, length(subsets))
seeds[[nCV*nRepeats+1]] <- sample.int(1000, 1)
```

### Random Forests

```{r}
rfeController <- rfeControl(functions   = rfFuncs,   # random forests
                            seeds       = seeds,
                            method      = "repeatedcv",
                            number      = nCV,          
                            repeats     = nRepeats,
                            verbose     = FALSE)
```

Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster) 

rfProfile <- rfe(x, y, sizes=subsets, rfeControl=rfeController)

stopCluster(rCluster)
```

```{r, comment=NA}
rfProfile
```

Output shows the best subset size was `r length(predictors(rfProfile))` predictors.

```{r, comment=NA}
predictors(rfProfile)
```

```{r, comment=NA
rfProfile$fit
```

```{r}
plot(rfProfile, type=c("g", "o"),
     main="rf:  Accuracy by Number of Variables")
```

```{r}
plot(rfProfile, type=c("g", "o"), metric="Kappa",
     main="rf:  Kappa by Number of Variables")
```

### Naive Bayes

```{r}
rfeController <- rfeControl(functions   = nbFuncs,   # naive bayes
                            seeds       = seeds,
                            method      = "repeatedcv",
                            number      = nCV,          
                            repeats     = nRepeats,
                            verbose     = FALSE)
```

Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster) 

nbProfile <- rfe(x, y, sizes=subsets, rfeControl=rfeController)

stopCluster(rCluster)
```

```{r, comment=NA}
nbProfile
```

Output shows the best subset size was `r length(predictors(nbProfile))` predictors.

```{r, comment=NA}
predictors(nbProfile)
```

```{r, comment=NA}
nbProfile$fit
```

```{r}
plot(nbProfile, type=c("g", "o"),
     main="nb:  Accuracy by Number of Variables")
```

```{r}
plot(nbProfile, type=c("g", "o"), metric="Kappa",
     main="nb:  Kappa by Number of Variables")
```

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`        
    
## References

[Feature Engineering using R](https://blogs.msdn.microsoft.com/microsoftrservertigerteam/2017/03/23/feature-engineering-using-r/), Microsoft, 2017. 

[Recursive Feature Elimination](http://ml-tutorials.kyrcha.info/rfe.html), Kyriakos Chatzidimitriou, 2017.