---
title: "Forensic Glass:  Boruta 'All Relevant' Variables"
output: 
  html_document:
    theme: united
    toc: yes
  html_notebook:
    theme: united
    toc: yes
---

efg | 2018-09-03

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
time.1 <- Sys.time()
```

[Boruta Package](https://cran.r-project.org/web/packages/Boruta/Boruta.pdf).  

*Finds relevant features by comparing original attributes' importance with importance achievable at random, estimated using their permuted copies (shadows).*

## Setup

```{r, comment=NA, message=FALSE, warning=FALSE}
library(tidyverse)
library(MASS)          # fgl data
library(caret)         # createDataParition
library(doParallel)    # registerDoParallel  
library(Boruta)        # Bortua
library(kableExtra)    
```

## Forensic Glass Data

```{r, comment=NA}
rawData <- fgl
dim(rawData)
```

```{r}
rawData                                           %>%
  kable("html", caption="Forensic Glass Data")    %>%
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left", font_size=12, 
                full_width=FALSE)                 %>%
  scroll_box(height="200px")
```

```{r, comment=NA}
table(rawData$type)
```

Relatively small "n" and class imbalance will make machine learning difficult.


## Define train and test datasets

```{r}
set.seed(71)

trainSetIndices <- createDataPartition(rawData$type, p=0.50, list=FALSE)

trainSet <- rawData[ trainSetIndices, ]
testSet  <- rawData[-trainSetIndices, ]   
```

```{r}
dim(trainSet)
```

## Boruta 'All Relevant' Variables

Setup parallel processing

```{r, comment=NA}
rCluster <- makePSOCKcluster(6)   # use 6 cores
registerDoParallel(rCluster)  
```

```{r, comment=NA}
set.seed(73)   # reproducible random numbers
BorutaModel <- Boruta(type ~ ., data=trainSet,
                      getImp = getImpFerns,
                      maxRuns = 1000, num.threads=6)
stopCluster(rCluster)
```

```{r, comment=NA}
print(BorutaModel)
```

```{r, comment=NA, fig.width=10, fig.height=8}
plot(BorutaModel, las=2, cex.axis=0.75, main="Boruta Importance")
grid()
```

```{r, comment=NA, fig.width=10, fig.height=8}
plotImpHistory(BorutaModel, main="Importance History")
grid()
```

```{r, comment=NA}
stats <- attStats(BorutaModel)
statsOrdered <- stats[order(stats$medianImp, decreasing=TRUE), ]

BorutaFeatures <- statsOrdered %>% rownames_to_column(var="Feature")

BorutaFeatures                                     %>%    
  kable("html", caption="Boruta Features")         %>%
  kable_styling(bootstrap_options=c("striped", "bordered", "condensed"),
                position="left", font_size=12, 
                full_width=FALSE)                  
```


```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`        
    
## References

* [Feature Selection in R with the Boruta R Package](https://www.datacamp.com/community/tutorials/feature-selection-R-boruta), DataCamp, 2018.  *Includes description of algorithm.*

* [Feature Selection: Select Important Variables using Boruta Algorithm](https://www.listendata.com/2017/05/feature-selection-boruta-package.html), Deepanshu Bhalla, 2017.  *Includes description of algorithm, comparison with Recursive Feature Elimination. *

* [Venn Diagram Comparison of Boruta, FSelectorRcpp and GLMnet Algorithms](http://r-addict.com/2016/06/19/Venn-Diagram-RTCGA-Feature-Selection.html), JMarcin Kosiński, 2016.

* [How to perform feature selection (i.e. pick important variables) using Boruta Package in R ?](https://www.analyticsvidhya.com/blog/2016/03/select-important-variables-boruta-package/), Analytics Vidhya, 2016.

* [Boruta - A System for Feature Selection](https://www.slideshare.net/brandonisherman/boruta-69779808), Miron B. Kursa, et al, 2016.

* [Feature selection: All-relevant selection with the Boruta package](https://www.r-bloggers.com/feature-selection-all-relevant-selection-with-the-boruta-package/), Allan Engelhardt, 2010.

* [Feature Selection with the Boruta Package](https://www.jstatsoft.org/article/view/v036i11), Miron B. Kursa and Witold R. Rudnicki, *Journal of Statistical Software*, 2010.

