---
title: "Forensic Glass:  Correlation Matrix"
output:
  html_document:
    theme: united
    toc: yes
  html_notebook:
    theme: united
    toc: yes
---

efg | 2018-09-03  

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
time.1 <- Sys.time()
```

## Removal of highly correlated variables 

Good idea for many machine learning algorithms based on linear algebra.  Not needed for treet based methods like random forest or Boruta.

Removal of constant variables is a good idea for efficiency.

## Setup

```{r, comment=NA, message=FALSE, warning=FALSE}
library(MASS)      # fgl data
library(corrplot)  # corrplot
library(caret)     # findCorrelation
library(tidyverse) # select
```

## Forensic Glass Data

```{r, comment=NA}
rawData <- fgl 
dim(rawData)
```

```{r}
rawData %>% head()
```

## Define train and test datasets

```{r}
set.seed(71)

trainSetIndices <- createDataPartition(rawData$type, p=0.75, list=FALSE)

trainSet <- rawData[ trainSetIndices, ]
testSet  <- rawData[-trainSetIndices, ]
```

```{r}
dim(trainSet)
```                    

## trainSet Correlation Matrix 

Use only training set data or we'll have a [data leak](https://machinelearningmastery.com/data-leakage-machine-learning/).

```{r, fig.width=8, fig.height=8}
colorScale <- colorRampPalette(c("#7F0000","red","#FF7F00","yellow","white",
                                 "cyan", "#007FFF", "blue","#00007F"))(100)  

corMatrix <- cor(trainSet %>% dplyr::select(-type))  # Create correlation matrix

corrplot(corMatrix, type="upper", method="square", order="AOE",
         tl.col="black", tl.srt=45, tl.cex=1.5,
         addCoef.col="darkgrey", number.cex=1.25,
         col=colorScale)
mtext("Correlation Matrix (angular order of eigenvectors)", line=3)         
```

## Removing highly correlated pairs with caret

Some machine learning algorithms are impeded by highly correlated predictors.  Caret's *findCorrelation* procedure can be used to remove one of the highly-correlated variables.

```{r, comment=NA}
HIGH_CORRELATION_CUTOFF <- 0.80

corHigh <- findCorrelation(corMatrix, HIGH_CORRELATION_CUTOFF)
if (length(corHigh) > 0)
{
  cat("Removing highly-correlated variable(s): ", names(trainSet)[corHigh])
  trainSet <- trainSet[, -corHigh] 
  testSet  <- testSet[, -corHigh]
}

trainSet %>% head()
```

In most machine learning applications, the *findCorrelation* check is applied to the training set data, but any highly-correlated predictors are also removed from the test set.

```{r, comment=NA, echo=FALSE}
time.2 <- Sys.time()
processingTime <- paste("Processing time:", sprintf("%.1f",
                        as.numeric(difftime(time.2, time.1, units="secs"))), "sec\n")
```

`r processingTime`
`r format(time.2, "%Y-%m-%d %H:%M:%S")`         

## References

* [corrplot: Visualization of a Correlation Matrix](https://cran.r-project.org/web/packages/corrplot/index.html), CRAN.
    
* [An Introduction to corrplot Package](https://cran.r-project.org/web/packages/corrplot/vignettes/corrplot-intro.html), CRAN.

* [Visualize correlation matrix using correlogram](http://www.sthda.com/english/wiki/visualize-correlation-matrix-using-correlogram), STHDA.

* [An Introduction to Matrix Visualization & corrplot Package](https://uploads.cosx.org/2009/12/An-Introduction-to-Matrix-Visualization-and-corrplot-Package.pdf), Taiyun Wei, 2009.

* [Introduction to Feature selection for bioinformaticians using R, correlation matrix filters, PCA & backward selection](Introduction to Feature selection for bioinformaticians using R, correlation matrix filters, PCA & backward selection), Yasset Perez-Riverol, 2013.